---
title: "Module 4 Project - Thera Bank"
author: "Mohanned Gomaa"
date: "2/15/2020"
output:
  word_document: default
  pdf_document: default
---

#Setting a working drict.
```{r}
setwd("C:/Users/mgomaa032/Desktop/PwC/Projects/DXB/PnO/WorkFroce analytics/UT Austin/Module 4/Week 3")
getwd()
```

#Importing master data.
```{r}
library(knitr)
library(readxl)
Bank_Full_DS= read_xlsx('Thera Bank_Personal_Loan_Modelling-dataset-1.xlsx', sheet = 'Bank_Personal_Loan_Modelling', col_names = T)
View(Bank_Full_DS)
tail(Bank_Full_DS,10)
```

#Expletory Analysis 
```{r}
dim(Bank_Full_DS)
names(Bank_Full_DS)
head(Bank_Full_DS)
View(Bank_Full_DS)
str(Bank_Full_DS)
summary(Bank_Full_DS)
```

#Missing Value Check
```{r}
sum(is.na(Bank_Full_DS))
colSums(is.na(Bank_Full_DS)) #18 missing values for family members.
```

#Treating Missing Values 
```{r}
Bank_Full_DS[is.na(Bank_Full_DS)]=0 #I took an assumption that NAs or empty cells under family members variable, represent single customers, so I replaced these empty cells with Zeros.
```

#Removing Unnecessary Columns
```{r}
Bank_Full_DS= Bank_Full_DS [,-1] #Removing Customer ID Column
```

#EDA - Basic data summary, Univariate, Bivariate analysis, graphs

```{r}
library(tidyverse)
library(dplyr)
library(corrplot)
corrplot(cor(Bank_Full_DS),method = 'square', title = 'Correlation matrix')

Loan_Customers.profile= Bank_Full_DS %>% group_by(Bank_Full_DS$`Personal Loan`)%>%
  summarise(Avg.Income= mean(Bank_Full_DS$`Income (in K/month)`),
            Avg.exp= mean(Bank_Full_DS$`Experience (in years)`), 
            Avg.CC= mean(Bank_Full_DS$CCAvg), 
            Med.CD = mode(Bank_Full_DS$`CD Account`))
print(Loan_Customers.profile)


hist(Bank_Full_DS$`Income (in K/month)`, main = "Income Profile",col='Navy', xlab = 'Income')
abline(v = mean(Bank_Full_DS$`Income (in K/month)`,col='Blue'))

hist(Bank_Full_DS$CCAvg, main = "Credit Card Profile",col='Navy', xlab = 'Avg Spend')
abline(v = mean(Bank_Full_DS$CCAvg,col='Blue'))

hist(Bank_Full_DS$`Experience (in years)`, main = "Experience Profile",col='Navy', xlab = 'Exp')
abline(v = mean(Bank_Full_DS$`Experience (in years)`,col='Blue'))

hist(Bank_Full_DS$Mortgage, main = "Mortgage Profile",col='Navy', xlab = 'Value')
abline(v = mean(Bank_Full_DS$Mortgage,col='Blue'))

boxplot(Bank_Full_DS$`Income (in K/month)`)
boxplot(Bank_Full_DS$CCAvg)
boxplot(Bank_Full_DS$`Experience (in years)`)
boxplot(Bank_Full_DS$Mortgage)
```

#Split Dataset into Train 70% and Test 30%
```{r}
library(caTools) #Call needed packages
set.seed(1000) #Setting a seed Conrtol Randomness 
spl= sample.split(Bank_Full_DS$`Personal Loan`,SplitRatio = 0.7) #Setting Split Ratio
Train_DS_Bank = subset(Bank_Full_DS,spl==T) #Train 70%
Test_DS_Bank = subset(Bank_Full_DS,spl== F) #Test 30%
dim(Train_DS_Bank) #Checking Dimensions 
dim(Test_DS_Bank) #Checking Dimensions
Test_DS_Bank_OG =subset(Bank_Full_DS,spl== F) 
head(Test_DS_Bank_OG)
view(Test_DS_Bank)
```


# 2.1 Clustering and 2.2 interpretation 
```{r}
#I’ve selected K-means (Centroid based) Clustering, as it is less complicated and expensive to its adversary, hierarchal clustering.  
head(Train_DS_Bank)
str(Train_DS_Bank)
Scale_Train_DS = scale(Train_DS_Bank) #Scale data to ensure that all columns has equal weights. 
str(Scale_Train_DS)
head(Scale_Train_DS) #view scaled Data
print(Scale_Train_DS) #view scaled Data
```

# K-means clustering before finetuning 
```{r}
set.seed (1000)
Clust_2 = kmeans(x=Scale_Train_DS,centers = 2, nstart = 5) #Setting up my clusters, at a random number of centers
print(Clust_2) #views results 
#Within cluster sum of squares by cluster:
  #[1] 27869.70 11780.18
  #(between_SS / total_SS =  12.8 %)
```

# plot K-means clustering
```{r}
library(cluster)
clusplot(Scale_Train_DS,Clust_2$cluster,color = T, shade = T ,lines =1) #The two clusters only explained ~33% of the point variability
```

# To improve model output, we need to find the right number for clusters
```{r}
#running 5 trials for no. of k with the least tot.withinss
totwss = rep(0,5)
for (k in 1:5) {set.seed(1000) 
  clust = kmeans(x=Scale_Train_DS,centers = k, nstart = 5)
  totwss[k] = clust$tot.withinss}
print(totwss)
plot (c(1:5),totwss, type = "b") #elbow method to identify the right number of clusters, plateau at 4
```

#Another way to find the right number of clusters, by through NbClust library
```{r}
library(NbClust)
set.seed(1000)
nc = NbClust(Scale_Train_DS,min.nc=2,max.nc=5,method ="kmeans")
```

```{r}
Clust_4 = kmeans(x=Scale_Train_DS,centers = 4, nstart = 5) 
print(Clust_4)
```

```{r}
Train_DS_Bank$Cluster= Clust_2$cluster
View(Train_DS_Bank)
custprofile = aggregate(Train_DS_Bank,list(Train_DS_Bank$Cluster), FUN = 'mean')
print(custprofile)
View(custprofile)
```

#Building Model 1 - CART 
#CART Decision Tree - CART 
```{r}
Train_DS_Bank_OG=Train_DS_Bank[,-14] 
View(Train_DS_Bank_OG)
head(Train_DS_Bank_OG)
nrow(Train_DS_Bank_OG)
sum(Train_DS_Bank_OG$`Personal Loan`=="1")/nrow(Train_DS_Bank_OG)*100 #Checking the percentage from my data who accepted the loan offer ~9.6%
```

#CART Install library
```{r}
library(rpart)
library(rpart.plot)
Train_DS_Bank_OG$`Personal Loan`= as.factor(Train_DS_Bank_OG$`Personal Loan`)
Test_DS_Bank_OG$`Personal Loan`= as.factor(Test_DS_Bank_OG$`Personal Loan`)
```

#CART Built the Tree
```{r}
CART= rpart(formula = Train_DS_Bank_OG$`Personal Loan`~.,
            data = Train_DS_Bank_OG, method = "class", minbucket = 3,cp = 0)
CART
```

#CART Plot and View the Tree
```{r}
rpart.plot(CART) #Resulted in a highly complex tree that needs to be pruned

printcp(CART)
plotcp(CART) # Cross Validation shows the lowest Error is at level 4, 5 splits 
```


#CART Prune to simplify the tree
```{r}
Pruned_CART = prune(CART, cp = 0.015,'CP')
printcp(Pruned_CART)
rpart.plot(Pruned_CART) #View the simplified Tree
Pruned_CART
```

#CART Busniess Rules 
```{r}
path.rpart(Pruned_CART,c(1:7))
```

#CART Prediction Outcome
```{r}
Pruned_CART_Out_Train=Train_DS_Bank_OG
Pruned_CART_Out_Train$Prediction = predict(Pruned_CART, data=Pruned_CART_Out_Train, type = "class")
Pruned_CART_Out_Train$Score = predict(Pruned_CART, data=Pruned_CART_Out_Train, type = "prob")[,'1']
head(Pruned_CART_Out_Train)
View(Pruned_CART_Out_Train)

library(dbplyr)
library(tidyverse)


Pruned_CART_Out_Test=Test_DS_Bank_OG

Pruned_CART_Out_Test$Prediction = Pruned_CART %>% predict(Pruned_CART_Out_Test, type = "class")

Pruned_CART_Out_Test$Score = (Pruned_CART %>% predict(Pruned_CART_Out_Test, type = "prob"))[,'1']

head(Pruned_CART_Out_Test)
View(Pruned_CART_Out_Train)
View(Pruned_CART_Out_Test)
```

#CART_Confusion Matrix - Model Performance 4.1
```{r}
CART_ConfusionMatrix_Train= table (Pruned_CART_Out_Train$`Personal Loan`,Pruned_CART_Out_Train$Prediction)

CART_ConfusionMatrix_Test= table (Pruned_CART_Out_Test$`Personal Loan`,Pruned_CART_Out_Test$Prediction)

print(CART_ConfusionMatrix_Train) #Very high accuracy
print(CART_ConfusionMatrix_Test) #Very high accuracy

print((CART_ConfusionMatrix_Train [1,2]+CART_ConfusionMatrix_Train [2,1]))/nrow(Pruned_CART_Out_Train) #Approx Error rate ~ 1.43%

print((CART_ConfusionMatrix_Test [1,2]+CART_ConfusionMatrix_Test [2,1]))/nrow(Pruned_CART_Out_Test) #Approx Error rate ~ 2.13%

print((CART_ConfusionMatrix_Train [1,1]+CART_ConfusionMatrix_Train [2,2]))/nrow(Pruned_CART_Out_Train) #Approx Accuracy rate ~ 98.6%

print((CART_ConfusionMatrix_Test [1,1]+CART_ConfusionMatrix_Test [2,2]))/nrow(Pruned_CART_Out_Test) #Approx Accuracy rate ~ 98%

print((CART_ConfusionMatrix_Train [2,2]))/(CART_ConfusionMatrix_Train [2,1]+CART_ConfusionMatrix_Train [2,2]) #Approx Precision rate ~ 87.5%

print((CART_ConfusionMatrix_Test [2,2]))/(CART_ConfusionMatrix_Test [2,1]+CART_ConfusionMatrix_Test [2,2]) #Approx Precision rate ~ 81%

print((CART_ConfusionMatrix_Train [2,2]))/(CART_ConfusionMatrix_Train [1,2]+CART_ConfusionMatrix_Train [2,2]) #Approx Sensitivity  rate ~ 97%

print((CART_ConfusionMatrix_Test [2,2]))/(CART_ConfusionMatrix_Test [1,2]+CART_ConfusionMatrix_Test [2,2]) #Approx Sensitivity  rate ~ 96.6%

print((CART_ConfusionMatrix_Train [1,1]))/(CART_ConfusionMatrix_Train [1,1]+CART_ConfusionMatrix_Train [2,1]) #Approx Specificity rate ~ 98.6%

print((CART_ConfusionMatrix_Test [1,1]))/(CART_ConfusionMatrix_Test [1,1]+CART_ConfusionMatrix_Test [2,1]) #Approx Specificity rate ~ 97.9%

```

#Break down my probs in Quantiles, as our focus to find the customers with the high prob to respond.
```{r}
CART_QS_Train = quantile(Pruned_CART_Out_Train$Score, prob= seq(0,1,length=11))

CART_QS_Test = quantile(Pruned_CART_Out_Test$Score, prob= seq(0,1,length=11))

print(CART_QS_Train)
print(CART_QS_Test)

plot(CART_QS_Train)
lines(CART_QS_Train)#Top 10 of all probs falls in 6th & and above

plot(CART_QS_Test)
lines(CART_QS_Test)#Top 10 of all probs falls in 6th & and above
```

#CART - Create a threshold for probs Cut-off
```{r}
CART_Threshold= CART_QS_Train [6]

mean(Pruned_CART_Out_Train$`Personal Loan`[Pruned_CART_Out_Train$Score > CART_Threshold]=='1') #Avg. Accuracy of my prediction is ~97.3%

mean(Pruned_CART_Out_Test$`Personal Loan`[Pruned_CART_Out_Test$Score > CART_Threshold]=='1') #Avg. Accuracy of my prediction is ~97%
```

#CART_Rank Order Table - Model Performance
```{r}
Pruned_CART_Out_Train$deciles = cut(Pruned_CART_Out_Train$Score,unique(CART_QS_Train),include.lowest = F)

Pruned_CART_Out_Test$deciles = cut(Pruned_CART_Out_Test$Score,unique(CART_QS_Test),include.lowest = F)

#Create deciles to bucket probabilities of each class.
print(Pruned_CART_Out_Train$deciles)
print(Pruned_CART_Out_Test$deciles)

head(Pruned_CART_Out_Train)
head(Pruned_CART_Out_Test)
```

#CART_Rank Order Table - Building Blocks 
```{r}
library(data.table)
CART_Train_DT = data.table(Pruned_CART_Out_Train)
CART_Test_DT = data.table(Pruned_CART_Out_Test)#Convert data in DT
View(CART_Train_DT)
CART_Train_rnk_tbl = CART_Train_DT[ ,list(cnt = length(CART_Train_DT$`Personal Loan`),cnt_tar1 =sum(CART_Train_DT$`Personal Loan`=="1"),cnt_tar0 =sum(CART_Train_DT$`Personal Loan`=="0")),by= deciles][order(-deciles)] #Base Count, Right and worng 
CART_Train_rnk_tbl$rrate= round(CART_Train_rnk_tbl$cnt_tar1/CART_Train_rnk_tbl$cnt)*100 # % Right
CART_Train_rnk_tbl$Cum_Resp = cumsum(CART_Train_rnk_tbl$cnt_tar1) #Cum right
CART_Train_rnk_tbl$Cum_Non_Resp = cumsum(CART_Train_rnk_tbl$cnt_tar0) #Cum wrong
CART_Train_rnk_tbl$Cum_rel_resp= round(CART_Train_rnk_tbl$Cum_Resp/sum(CART_Train_rnk_tbl$cnt_tar1),4)*100 # % Cum right
CART_Train_rnk_tbl$Cum_rel_Non_resp= round(CART_Train_rnk_tbl$Cum_Non_Resp/sum(CART_Train_rnk_tbl$cnt_tar0),4)*100 # % Cum wrong
CART_Train_rnk_tbl$Ks= abs(CART_Train_rnk_tbl$Cum_rel_resp-CART_Train_rnk_tbl$Cum_rel_Non_resp) # Ks

CART_Test_rnk_tbl = CART_Test_DT[ ,list(cnt = length(CART_Test_DT$`Personal Loan`),cnt_tar1 =sum(CART_Test_DT$`Personal Loan`=="1"),cnt_tar0 =sum(CART_Test_DT$`Personal Loan`=="0")),by= deciles][order(-deciles)] #Base Count, Right and worng 
CART_Test_rnk_tbl$rrate= round(CART_Test_rnk_tbl$cnt_tar1/CART_Test_rnk_tbl$cnt)*100 # % Right
CART_Test_rnk_tbl$Cum_Resp = cumsum(CART_Test_rnk_tbl$cnt_tar1) #Cum right
CART_Test_rnk_tbl$Cum_Non_Resp = cumsum(CART_Test_rnk_tbl$cnt_tar0) #Cum wrong
CART_Test_rnk_tbl$Cum_rel_resp= round(CART_Test_rnk_tbl$Cum_Resp/sum(CART_Test_rnk_tbl$cnt_tar1),4)*100 # % Cum right
CART_Test_rnk_tbl$Cum_rel_Non_resp= round(CART_Test_rnk_tbl$Cum_Non_Resp/sum(CART_Test_rnk_tbl$cnt_tar0),4)*100 # % Cum wrong
CART_Test_rnk_tbl$Ks= abs(CART_Test_rnk_tbl$Cum_rel_resp-CART_Test_rnk_tbl$Cum_rel_Non_resp) # Ks

print(CART_Train_rnk_tbl)
print(CART_Test_rnk_tbl)
```

#CART_ ROC, AUC and Gini - Model performance
```{r}
library(ROCR)
library(ineq)
CART_pred_Obj_Train = prediction(Pruned_CART_Out_Train$Score, Pruned_CART_Out_Train$`Personal Loan`) # Building ROC and  compare
CART_pref_Train = performance(CART_pred_Obj_Train, "tpr","fpr") # Calculate TPR and FPR
plot(CART_pref_Train) # Plot performance curve 

CART_pred_Obj_Test = prediction(Pruned_CART_Out_Test$Score, Pruned_CART_Out_Test$`Personal Loan`) # Building ROC and  compare
CART_pref_Test = performance(CART_pred_Obj_Test, "tpr","fpr") # Calculate TPR and FPR
plot(CART_pref_Test) # Plot performance curve 

#Calculate KS
CART_Train_Ks = max(CART_pref_Train@y.values[[1]]-CART_pref_Train@x.values[[1]])
print(CART_Train_Ks) #K is 91.2 

CART_Test_Ks = max(CART_pref_Test@y.values[[1]]-CART_pref_Test@x.values[[1]])
print(CART_Test_Ks) #K is 92.4

#Calculate AUC
CART_AUC_Train = performance(CART_pred_Obj_Train,"auc")
CART_AUC_Train = as.numeric(CART_AUC_Train@y.values)
print (CART_AUC_Train) # AUC is ~98

CART_AUC_Test = performance(CART_pred_Obj_Test,"auc")
CART_AUC_Test = as.numeric(CART_AUC_Test@y.values)
print (CART_AUC_Test) # AUC is ~98

#Calculate Gini
CART_Gini_Train = ineq(Pruned_CART_Out_Train$Score,"gini")
CART_Gini_Test = ineq(Pruned_CART_Out_Test$Score,"gini")
print(CART_Gini_Train) # Gini equal 0.871
print(CART_Gini_Test) # Gini equal 0.874
```

#CART_ Concordance and Discordance ratios'
```{r}
library(InformationValue)
Concordance(actuals = Pruned_CART_Out_Train$`Personal Loan`, predictedScores = Pruned_CART_Out_Train$Score)

#Concordance 0.9650382
#Discordance 0.03496177
#Tied 4.857226e-17

Concordance(actuals = Pruned_CART_Out_Test$`Personal Loan`, predictedScores = Pruned_CART_Out_Test$Score)

#Concordance 0.9670293
#Discordance 0.03297075
#Tied -4.163336e-17
```


##Building Random Forest Model 
#Random Forest Install/Call Library
```{r}
sum(Train_DS_Bank_OG$Personal.Loan=="1")
sum(Train_DS_Bank_OG$Personal.Loan=="1")/nrow(Train_DS_Bank_OG)*100 #Checking the percentage from my data who accepted the loan offer
library(randomForest)
library(caret)#Build Confusion Matrix 
library(e1071)
```

#Build Random Forest
```{r}
head(Train_DS_Bank_OG)
set.seed(1000)
names(Train_DS_Bank_OG) = make.names(names(Train_DS_Bank_OG))
names(Test_DS_Bank) = make.names(names(Test_DS_Bank))

RndForest = randomForest(Train_DS_Bank_OG$Personal.Loan~.,data = Train_DS_Bank_OG,ntree=501,mtry=3,nodesize=10,importance=T)
print(RndForest)
```

#Evaluate Error rate and try to reduce it
```{r}
print(RndForest$err.rate)
plot(RndForest) #Based on the plot we see that OOB plateau after ~51, so no point to use 501 trees in our forest
```
#Evaluate the important of the variables used on our model
```{r}
importance(RndForest) #The most important two variables are Income and Education for our predication 
View(Train_DS_Bank_OG)
```

#Run number of Scenarios for mtry or variables randomly selected to understand what is the impact on my OOB.
```{r}
set.seed(1000)
T_RndForest = tuneRF(Train_DS_Bank_OG[,-9],Train_DS_Bank_OG$Personal.Loan,mtryStart = 6,stepFactor = 1.5,ntreeTry = 51,improve = 0.0001, nodesize = 10,trace = T,plot = T,doBest = T,importance=T) #6 Variables is the best, yielding the lowest OOB
```

#Testing my RF model predication Power
```{r}
RF_OutPut_Train = Train_DS_Bank_OG
RF_OutPut_Test = Test_DS_Bank

RF_OutPut_Train$Predict.Class = predict(T_RndForest,RF_OutPut_Train,type = "class")

RF_OutPut_Test$Predict.Class = predict(T_RndForest,RF_OutPut_Test,type = "class")

RF_OutPut_Train$Predict.Score = predict(T_RndForest,RF_OutPut_Train,type = "prob")[,"1"] 

RF_OutPut_Test$Predict.Score = predict(T_RndForest,RF_OutPut_Test,type = "prob")[,"1"] 
#Porb for who expected the personal loan offer
View(RF_OutPut_Train)
View(RF_OutPut_Test)
```

#RF_Confusion Matrix - Model Performance 4.1
```{r}
RF_ConfusionMatrix_Train= table (RF_OutPut_Train$Personal.Loan,RF_OutPut_Train$Predict.Class)

library(caret)
confusionMatrix(RF_OutPut_Train$Personal.Loan,RF_OutPut_Train$Predict.Class)

RF_ConfusionMatrix_Test= table (RF_OutPut_Test$Personal.Loan,RF_OutPut_Test$Predict.Class)

print(RF_ConfusionMatrix_Train) #Very high accuracy
print(RF_ConfusionMatrix_Test) #Very high accuracy

print((RF_ConfusionMatrix_Train [1,2]+RF_ConfusionMatrix_Train [2,1]))/nrow(RF_OutPut_Train) #Approx Error rate ~ 0.66%

print((RF_ConfusionMatrix_Test [1,2]+RF_ConfusionMatrix_Test [2,1]))/nrow(RF_OutPut_Test) #Approx Error rate ~ 1.9%

print((RF_ConfusionMatrix_Train [1,1]+RF_ConfusionMatrix_Train [2,2]))/nrow(RF_OutPut_Train) #Approx Accuracy rate ~ 99%

print((RF_ConfusionMatrix_Test [1,1]+RF_ConfusionMatrix_Test [2,2]))/nrow(RF_OutPut_Test) #Approx Accuracy rate ~ 98%

print((RF_ConfusionMatrix_Train [2,2]))/(RF_ConfusionMatrix_Train [2,1]+RF_ConfusionMatrix_Train [2,2]) #Approx Precision rate ~ 93.5%

print((RF_ConfusionMatrix_Test [2,2]))/(RF_ConfusionMatrix_Test [2,1]+RF_ConfusionMatrix_Test [2,2]) #Approx Precision rate ~ 83.3%

print((RF_ConfusionMatrix_Train [2,2]))/(RF_ConfusionMatrix_Train [1,2]+RF_ConfusionMatrix_Train [2,2]) #Approx Sensitivity  rate ~ 100%

print((RF_ConfusionMatrix_Test [2,2]))/(RF_ConfusionMatrix_Test [1,2]+RF_ConfusionMatrix_Test [2,2]) #Approx Sensitivity  rate ~ 97%

print((RF_ConfusionMatrix_Train [1,1]))/(RF_ConfusionMatrix_Train [1,1]+RF_ConfusionMatrix_Train [2,1]) #Approx Specificity rate ~ 99%

print((RF_ConfusionMatrix_Test [1,1]))/(RF_ConfusionMatrix_Test [1,1]+RF_ConfusionMatrix_Test [2,1]) #Approx Specificity rate ~ 98%
```

#Break down my probs in Quantiles, as our focus to find the customers with the high prob to respond.
```{r}
RF_QS_Train = quantile(RF_OutPut_Train$Predict.Score, prob= seq(0,1,length=11))

RF_QS_Test = quantile(RF_OutPut_Test$Predict.Score, prob= seq(0,1,length=11))

print(RF_QS_Train)
print(RF_QS_Test)

plot(RF_QS_Train)
lines(RF_QS_Train)#Top 10 of all probs falls in 10th & 11th range

plot(RF_QS_Test)
lines(RF_QS_Test)#Top 10 of all probs falls in 10th & 11th range
```

#Create a threshold for probs Cut-off
```{r}
RF_Threshold= RF_QS_Train [10]

mean(RF_OutPut_Train$Personal.Loan[RF_OutPut_Train$Predict.Score>RF_Threshold]=="1") #Avg. Accuracy of my prediction is ~96%

mean(RF_OutPut_Test$Personal.Loan[RF_OutPut_Test$Predict.Score>RF_Threshold]=="1") #Avg. Accuracy of my prediction is ~95%
```


#RF_Rank Order Table - Model Performance
```{r}
RF_OutPut_Train$deciles = cut(RF_OutPut_Train$Predict.Score,unique(RF_QS_Train), inculde.lowest = F)
RF_OutPut_Test$deciles = cut(RF_OutPut_Test$Predict.Score,unique(RF_QS_Test), inculde.lowest = F) #Create deciles to bucket probabilities of each class.

print(RF_OutPut_Train$deciles)
print(RF_OutPut_Test$deciles)

head(RF_OutPut_Train)
head(RF_OutPut_Test)
```

#RF_Rank Order Table - Building Blocks 
```{r}
library(data.table)
RF_Train_DT = data.table(RF_OutPut_Train)
RF_Test_DT = data.table(RF_OutPut_Test)#Convert data in DT
View(RF_Train_DT)
RF_Train_rnk_tbl = RF_Train_DT[ ,list(cnt = length(RF_Train_DT$Personal.Loan),cnt_tar1 =sum(RF_Train_DT$Personal.Loan=="1"),cnt_tar0 =sum(RF_Train_DT$Personal.Loan=="0")),by= deciles][order(-deciles)] #Base Count, Right and worng 
RF_Train_rnk_tbl$rrate= round(RF_Train_rnk_tbl$cnt_tar1/RF_Train_rnk_tbl$cnt)*100 # % Right
RF_Train_rnk_tbl$Cum_Resp = cumsum(RF_Train_rnk_tbl$cnt_tar1) #Cum right
RF_Train_rnk_tbl$Cum_Non_Resp = cumsum(RF_Train_rnk_tbl$cnt_tar0) #Cum wrong
RF_Train_rnk_tbl$Cum_rel_resp= round(RF_Train_rnk_tbl$Cum_Resp/sum(RF_Train_rnk_tbl$cnt_tar1),4)*100 # % Cum right
RF_Train_rnk_tbl$Cum_rel_Non_resp= round(RF_Train_rnk_tbl$Cum_Non_Resp/sum(RF_Train_rnk_tbl$cnt_tar0),4)*100 # % Cum wrong
RF_Train_rnk_tbl$Ks= abs(RF_Train_rnk_tbl$Cum_rel_resp-RF_Train_rnk_tbl$Cum_rel_Non_resp) # Ks
print(RF_Train_rnk_tbl)

RF_Test_rnk_tbl = RF_Test_DT[ ,list(cnt = length(RF_Test_DT$Personal.Loan),cnt_tar1 =sum(RF_Test_DT$Personal.Loan=="1"),cnt_tar0 =sum(RF_Test_DT$Personal.Loan=="0")),by=deciles][order(-deciles)]
RF_Test_rnk_tbl$rrate= round(RF_Test_rnk_tbl$cnt_tar1/RF_Test_rnk_tbl$cnt)*100
RF_Test_rnk_tbl$Cum_Resp = cumsum(RF_Test_rnk_tbl$cnt_tar1)
RF_Test_rnk_tbl$Cum_Non_Resp = cumsum(RF_Test_rnk_tbl$cnt_tar0)
RF_Test_rnk_tbl$Cum_rel_resp= round(RF_Test_rnk_tbl$Cum_Resp/sum(RF_Test_rnk_tbl$cnt_tar1),4)*100
RF_Test_rnk_tbl$Cum_rel_Non_resp= round(RF_Test_rnk_tbl$Cum_Non_Resp/sum(RF_Test_rnk_tbl$cnt_tar0),4)*100
RF_Test_rnk_tbl$Ks= abs(RF_Test_rnk_tbl$Cum_rel_resp-RF_Test_rnk_tbl$Cum_rel_Non_resp)
print(RF_Test_rnk_tbl)
```

#RF_ ROC, AUC and Gini - Model performance
```{r}
library(ROCR)
library(ineq)
RF_pred_Obj_Train = prediction(RF_OutPut_Train$Predict.Score, RF_OutPut_Train$Personal.Loan) # Building ROC and  compare
RF_pref_Train = performance(RF_pred_Obj_Train, "tpr","fpr") # Calculate TPR and FPR
plot(RF_pref_Train) # Plot performance curve 

RF_pred_Obj_Test = prediction(RF_OutPut_Test$Predict.Score, RF_OutPut_Test$Personal.Loan) # Building ROC and  compare
RF_pref_Test = performance(RF_pred_Obj_Test, "tpr","fpr") # Calculate TPR and FPR
plot(RF_pref_Test) # Plot performance curve 

#Calculate KS
RF_Train_Ks = max(RF_pref_Train@y.values[[1]]-RF_pref_Train@x.values[[1]])
print(RF_Train_Ks) #K is 99.5

RF_Test_Ks = max(RF_pref_Test@y.values[[1]]-RF_pref_Test@x.values[[1]])
print(RF_Test_Ks) #K is 96.8 

#Calculate AUC
RF_AUC_Train = performance(RF_pred_Obj_Train,"auc")
RF_AUC_Train = as.numeric(RF_AUC_Train@y.values)
print (RF_AUC_Train) # AUC is ~100

RF_AUC_Test = performance(RF_pred_Obj_Test,"auc")
RF_AUC_Test = as.numeric(RF_AUC_Test@y.values)
print (RF_AUC_Test) # AUC is ~99.8

#Calculate Gini
RF_Gini_Train = ineq(RF_OutPut_Train$Predict.Score,"gini")
RF_Gini_Test = ineq(RF_OutPut_Test$Predict.Score,"gini")
print(RF_Gini_Train) # Gini equal 0.898
print(RF_Gini_Test) # Gini equal 0.898
```

#RF_ Concordance and Discordance ratios'
```{r}
library(InformationValue)
Concordance(actuals = RF_OutPut_Train$Personal.Loan, predictedScores = RF_OutPut_Train$Predict.Score)

#Concordance 0.9998504
#Discordance 0.000149562
#Tied 2.493665e-17
Concordance(actuals = RF_OutPut_Test$Personal.Loan, predictedScores = RF_OutPut_Test$Predict.Score)
#Concordance 0.9977927
#Discordance 0.002207268
#Tied -7.806256e-18
```



